Deploying and Sharing QSAR Models using Predictive Modeling Markup Language

### 2014-12-05 paper.tex
Mathematical relationships in QSAR

>>>>

\subsection{Common mathematical relationships in QSAR}
Three major types:
\begin{itemize}
  \item Standalone. The output is computed by applying a function to the input:
  \begin{equation}
    output = f(input)
  \end{equation}
  Most relationships can be efficiently represented using top-level PMML model
  elements (15 elements as of PMML schema version 4.2). Some relationships (eg. PCA,
  PLS) are not distinctive enough to warrant dedicated PMML model elements.
  They are represented by combining some analogous PMML model element with appropriate 
  data pre- or post-processing transformations. For example, a PLS model can be
  represented by a combination of a RegressionModel element with NormContinuous 
  data transformation elements (ie. one for each regressor).
  \item Model aggregation. The final output is computed by applying some aggregation
  function over a collection of interim outputs:
  \begin{equation}
    output = f\_aggregate(f\_1(input), f\_2(input), ..)
  \end{equation}
  All primary functions must be of the same type. Most common aggregation functions 
  are the average and weighted average for regression, and the majority vote and weighted 
  majority vote for classification and clustering. Starting from PMML schema version
  4.0, such model aggregates can be very efficiently represented using the segmentation
  approach (eg. a MiningModel element with a Segmentation child element).
  Model aggregation encompasses wide range of boosting and bagging models [ref] that
  have been gaining ground since mid 2000s and often demonstrate superior predictive
  capabilities. For example, the Random Forest model does not warrant a dedicated PMML
  element, because it can be completely represented as an aggregation of any number
  (typically, in the hundreds) of decision tree models.
  Main QSAR application is related to consensus modeling. Aggregated models can be 
  of same type (eg. a collection of linear regression models) or different types 
  (eg. a mix of linear regression, neural network regression and SVM regression models).
  Such consensus models have 
  \item Model chain. The final output is computed by applying a function to the
  the interim output:
  \begin{equation}
    output = f\_2(f\_1(input))
  \end{equation}
  Primary functions can be of any type. Typically, a classification model chained to
  several class-based regression models. Again, represented using the segmentation
  approach. The output of one model may need some transformation in order to be
  usable as input for the subsequent model.
  Main QSAR application is mechanistic models. For example, consider the a workflow 
  for predicting the acute aquatic toxicity of small organic molecules. First, the
  incoming chemical compound is assigned to some class based on its predicted mechanism
  of toxic action. Second, the the class label is used to select the corresponding
  regression model for predicting the effective concentration (eg. pLC50, pIGC50).
  PMML provides facilities for encapsulating model chains as a single predictive 
  solution. From end-user perspective, much more convenient to obtain the final output
  in single function call than to emulate it manually by evaluating multiple 
  standalone models.
\end{itemize}
